State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-06-14 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
more shorten

Figure \ref{fg_robustness_scores} presents the scores for surface robustness.
The packages with higher scores elegantly handled unexpected/unanticipated
inputs, typically showing a clear error message. We may have underscored
\textit{OHIF Viewer}, since we needed further customization to load
data.

@@

According to their documentation, all 29 software packages should support the
DICOM standard. To test robustness, we prepared two types of image files:
correct and incorrect formats (with the incorrect format created by relabelled a
text file to have the ``.dcm'' extension).  All software packages loaded the
correct format image, except for \textit{GATE}, which failed for unknown
reasons.  For the broken format, \textit{MatrixUser}, \textit{dwv}, and
\textit{Slice:Drop} ignored the incorrect format of the file and loaded it
regardless. They did not show any error message and displayed a blank image.
\textit{MRIcroGL} behaved similarly except that it showed a meaningless image.
\textit{Drishti} successfully detected the broken format of the file, but the
software crashed as a result.

@@

\begin{figure}[!ht]
\includegraphics[scale=0.38]{usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

@@

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. Only a few of the 29 projects had a
product, developer's manual, or API documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents. Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%).
Table~\ref{tab_maintainability_docs} shows which projects had these documents,
in descending order of their maintainability scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lccc}
\toprule
\multicolumn{1}{c}{Software} & Prod.\ Roadmap & Dev.\ Manual & API Doc. \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark \\
Weasis &  & \checkmark &  \\
OHIF Viewer &  & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark \\
ParaView & \checkmark &  &  \\
SMILI &  &  & \checkmark \\
medInria &  & \checkmark &  \\
INVESALIUS 3 & \checkmark &  &  \\
dwv &  &  & \checkmark \\
BioImage Suite Web &  & \checkmark &  \\
Gwyddion &  & \checkmark & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with the maintainability documents (listed in descending order of 
maintainability score)}
\label{tab_maintainability_docs}
\end{table}

Twenty-seven of the 29 projects used git as the version control tool, with 24 of
these using GitHub. \textit{AMIDE} used Mercurial and \textit{Gwyddion} used
Subversion. \textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used
SourceForge. \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

@@

Figure~\ref{fg_reusability_scores} shows the AHP results for reusability. As
described in Section~\ref{sec_grading_software}, we gave higher scores to the
projects with API documentation. As shown in
Table~\ref{tab_maintainability_docs}, seven projects had API documents. We also
assumed that projects with more code files and less LOC per code file are more
reusable. Table \ref{tab_loc_per_file} shows the number of text-based files by
project, which we used to approximate the number of code files. The table also
lists the total number of lines (including comments and blanks), LOC, and
average LOC per file. We arranged the items in descending order of their
reusability scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (sorted in descending order of reusability
scores)}
\label{tab_loc_per_file}
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.38]{understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

@@

Figure~\ref{fg_visibility_transparency_scores} shows the AHP scores for
visibility/transparency. Generally speaking, the teams that actively documented
their development process and plans scored higher.
Table~\ref{tab_Visibility/Transparency_docs} shows the projects that had
documents for the development process, project status, development environment,
and release notes, in descending order of their visibility/transparency
scores.

\begin{figure}[!ht]
\includegraphics[scale=0.38]{visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

@@

As described in Section~\ref{sec_AHP}, for our AHP measurements, we have nine
criteria (qualities) and 29 alternatives (software packages). In the absence of
a specific real world context, we assumed all nine qualities are equally
important. Figure~\ref{fg_overall_scores} shows the overall scores in descending
order. Since we produced the scores from the AHP process, the total sum of the
29 scores is precisely 1.0.

>>>>

Figure \ref{fg_robustness_scores} presents the scores for surface robustness.
The packages with higher scores elegantly handled unexpected or unanticipated
inputs, typically showing a clear error message. We may have underscored
\textit{OHIF Viewer}, since we needed further customization to load
data.

@@

According to their documentation, all 29 software packages should support the
DICOM standard. To test robustness, we prepared two types of image files:
correct and incorrect formats (with the incorrect format created by relabelling a
text file to have the ``.dcm'' extension).  All software packages loaded the
correct format image, except for \textit{GATE}, which failed for unknown
reasons.  For the broken format, \textit{MatrixUser}, \textit{dwv}, and
\textit{Slice:Drop} ignored the incorrect format,
did not show any error message and displayed a blank image.
\textit{MRIcroGL} behaved similarly except that it showed a meaningless image.
\textit{Drishti} successfully detected the broken format of the file, but the
software crashed as a result.

@@

\begin{figure}[ht]
\includegraphics[scale=0.38]{usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

@@

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. Only a few of the 29 projects had a
product, developer's manual, or API documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents 
(see Table~\ref{tab_maintainability_docs} the full data).  Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%).

\begin{figure}[ht]
\includegraphics[scale=0.38]{maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{lccc}
\toprule
\multicolumn{1}{c}{Software} & Prod.\ Roadmap & Dev.\ Manual & API Doc. \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark \\
Weasis &  & \checkmark &  \\
OHIF Viewer &  & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark \\
ParaView & \checkmark &  &  \\
SMILI &  &  & \checkmark \\
medInria &  & \checkmark &  \\
INVESALIUS 3 & \checkmark &  &  \\
dwv &  &  & \checkmark \\
BioImage Suite Web &  & \checkmark &  \\
Gwyddion &  & \checkmark & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with the maintainability documents (listed in descending order of 
maintainability score)}
\label{tab_maintainability_docs}
\end{table}

Twenty-seven of the 29 projects used git for version control, with 24 of
these using GitHub. \textit{AMIDE} used Mercurial and \textit{Gwyddion} used
Subversion. \textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used
SourceForge. \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

@@

API documents (see Table~\ref{tab_maintainability_docs}) also help maintainability,
which explains some of the scores shown in
Figure~\ref{fg_reusability_scores}. We have assumed that smaller code files are
likely more reusable -- see
Table \ref{tab_loc_per_file} for the details.

\begin{figure}[ht]
\includegraphics[scale=0.38]{reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (by reusability scores)}
\label{tab_loc_per_file}
\end{table}

@@

\begin{figure}[ht]
\includegraphics[scale=0.38]{understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

@@

Figure~\ref{fg_visibility_transparency_scores} shows the AHP scores for
visibility/transparency. Generally speaking, the teams that actively documented
their development process and plans scored higher.
Table~\ref{tab_Visibility/Transparency_docs} shows the projects that had
documents for the development process, project status, development environment,
and release notes.

\begin{figure}[ht]
\includegraphics[scale=0.38]{visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

@@

In the absence of
a specific real world context, we assumed all nine qualities are equally
important. Figure~\ref{fg_overall_scores} shows the overall scores in descending
order.
